An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-14 draft.tex
Motivations

\begin{abstract} 
Multi-population based methods have been used extensively in
recent years to improve the performance of nature-inspired optimization
algorithms. These methods are parallel and asynchronous by nature and researchers
leverage this to scale implementations 
applying distributed, multi-threaded, parallel, and cloud-based computing. 
% scale implementations with ... implementations. Please rephrase - JJ 
% done - Mario
However, the design  of these cloud-native,  
distributed, multi-population algorithms is not a trivial task. Designers need
to adapt their solutions, 
% design of algorithms is not trivial... designer need to adapt the algorithms... Maybe rephrase to avoid repetition - JJ to leverage
the scalability, elasticity, fault-tolerance, reproducibility, and
cost-effectiveness of cloud systems while conserving the intended functionality.
% I am removing the native part above, because is also very repetetive - Mario 
% Something about "this is the problem we want to address... " to link the previous sentence with this one - JJ 
Furthermore, currently, there are no
cloud-native solutions that support basic algorithm design practices that can make
multi-population methods more efficient: having multiple search strategies and
varying the number and size of subpopulations. We address these problems in this work,
by presenting the design and implementation of a cloud-native 
container-based % again maybe it's better cloud-native instead of reactive - JJ
application for the execution of ensemble multi-population algorithms. Researchers can
use the application to execute multiple experiments, both locally or in a cloud
environment and deploy experiments by specifying the infrastructure as part of
an experiment definition so that results are easily reproducible.
% A sentence on how this implementation is better than the state of the art - JJ 
% Added a comment above - Mario, 
% But maybe we need to say it again? - Mario
% For instance "Generally the current implementations include a single algorithm, which is bad because... " - JJ 
% Also "Current implementations are more or less functional equivalent 
%to parallel, including the fact that they run copies of the same, parametrized, algorithm... 
% If this is too long, maybe just some reference about 
%"current implementation do not leverage the fact that (you can use different algorithms,
% and differently parametrized...) and then qualify this in the introduction - JJ
State of the art approaches use a single-algorithm implementation, and this might
not be the best solution. We compared the speedup, execution time, and the number
of evaluations between a single and an ensemble approach by using Genetic Algorithms (GAs)
and Particle Swarm Optimization (PSO) algorithms against a benchmark for the optimization
of continuous functions. Experiments show that the framework allows the ensemble 
to outperform single-algorithm versions,  
% We also run experiments to see the effect of population size for seveal problem sizes  
% Against NUMBER OF WORKERS (this concept is not mentioned yet)
% We can add something like:      
Moreover, we evaluated the effect other design parameters had on the
performance; in particular, the number and the size of populations concerning
the problem size. The architecture and the implemented platform is an excellent
alternative for locally or in the cloud, thus proving that cloud-native
bioinspired algorithms perform better in their "natural" environment than other
kinds of algorithms.
\end{abstract}

@@

Taking these factors into account, in this work, we present EcoSwarm, % maybe a less generic name? - JJ
% agree - Mario
a multi-container Docker application that reactively processes isolated and
heterogeneous subpopulations. The platform uses the Docker Compose tool for
defining and running experiments as multi-container applications. 
% Docker compose is not really for production, more for development environments, 
% you have to justify this. Also, it does not handle scaling so need to justify it too - JJ
Docker Container images host the stateless functions responsible for running algorithms
on subpopulations. Researchers then define the configuration of services used by
a multi-population algorithm using a YAML file. Any researcher can deploy and
start all the services from this configuration file with a single command. Once
the services are up, multiple instances of an experiment can be sent to a queue
for their execution. The platform includes containerized services that are
compatible with other container orchestration technologies like Kubernetes and
Docker Swarm. The services included are a Redis Server used for the Message
Queue implementation and logging and an Experiment Controller implemented in
Python. Developers of sub-population processing functions have the freedom to
implement new population-based algorithms or operators in the language they
choose inside a container, as they only need to subscribe to a channel of the
Message Queue and have the ability to read the population and parameters as JSON  
file.

>>>>

\begin{abstract} 
Multi-population based methods have been used extensively in
recent years to improve the performance of nature-inspired optimization
algorithms. These methods are parallel and asynchronous by nature and researchers
leverage this to scale implementations 
applying distributed, multi-threaded, parallel, and cloud-based computing. 
% scale implementations with ... implementations. Please rephrase - JJ 
% done - Mario
However, the design  of these cloud-native,  
distributed, multi-population algorithms is not a trivial task. Designers need
to adapt their solutions 
% design of algorithms is not trivial... designer need to adapt the algorithms... Maybe rephrase to avoid repetition - JJ 
to leverage
the scalability, elasticity, fault-tolerance, reproducibility, and
cost-effectiveness of cloud systems while conserving the intended functionality.
% I am removing the native part above, because is also very repetetive - Mario 
% Something about "this is the problem we want to address... " to link the previous sentence with this one - JJ 
Furthermore, currently, there are no
cloud-native solutions that support basic algorithm design practices that can make
multi-population methods more efficient: having multiple search strategies, and
varying the number and size of subpopulations. We address these problems in this work,
by presenting the design and implementation of a cloud-native 
container-based % again maybe it's better cloud-native instead of reactive - JJ
application for the execution of ensemble multi-population algorithms. Researchers can
use the application to execute multiple experiments, both locally or in a cloud
environment, and deploy experiments by specifying the infrastructure as part of
an experiment definition so that results are easily reproducible.
% A sentence on how this implementation is better than the state of the art - JJ 
% Added a comment above - Mario, 
% But maybe we need to say it again? - Mario
% For instance "Generally the current implementations include a single algorithm, which is bad because... " - JJ 
% Also "Current implementations are more or less functional equivalent 
%to parallel, including the fact that they run copies of the same, parametrized, algorithm... 
% If this is too long, maybe just some reference about 
%"current implementation do not leverage the fact that (you can use different algorithms,
% and differently parametrized...) and then qualify this in the introduction - JJ
State of the art approaches use a single-algorithm implementation, and this might
not be the best solution. We compared the speedup, execution time, and the number
of evaluations between a single and an ensemble approach by using Genetic Algorithms (GAs)
and Particle Swarm Optimization (PSO) algorithms against a benchmark for the optimization
of continuous functions. Experiments show that the framework allows the ensemble 
to outperform single-algorithm versions.  
% We also run experiments to see the effect of population size for seveal problem sizes  
% Against NUMBER OF WORKERS (this concept is not mentioned yet)
% We can add something like:      
Moreover, we evaluated the effect other design parameters had on the
performance, in particular, the number and the size of populations concerning
the problem size. The architecture and the implemented platform is an excellent
alternative for locally or in the cloud, thus proving that cloud-native
bioinspired algorithms perform better in their "natural" environment than other
kinds of algorithms.
\end{abstract}

@@

Researchers and algorithm designers willing to  adapt to the cloud their
multi-population solutions face the following challenges:
\begin{itemize}
    \item They need to change their current solutions to a reactive architecture in which processes
    communicate with each other by exchanging messages asynchronously and react
    to a continuous stream of data. One of the options is to consider
    sub-populations as messages to be modified asynchronously by functions.

    \item To scale the system, when it is possible, they must use stateless functions for
    processing.

    \item They need to establish a workflow of local development and prototyping while having the 
    option of deployment in a cloud provider. It is desirable to use modern development methodologies and technologies.

    \item They must log experimental results and be able to replicate the experiments.

    \item They must consider additional parameters that can affect the monetary cost or the performance of the
    execution — for instance, the format and size of messages, the number of
    worker processes and their capabilities.  
\end{itemize}

Taking these factors into account, in this work, we present EcoSwarm, % maybe a less generic name? - JJ
% agree - Mario
a multi-container Docker application that reactively processes isolated and
heterogeneous subpopulations. The platform uses the Docker Compose tool for
defining and running experiments as multi-container applications. 
% Docker compose is not really for production, more for development environments, 
% you have to justify this. Also, it does not handle scaling so need to justify it too - JJ
% justification: - Mario
We preferred the use of a container-based application for this prototype because
it is more suitable for development and replication without the need for
additional components. Nevertheless, the application is compatible with other
orchestration and deployment technologies if more scalabilty like Docker Swarm
and Kubernetes with additional configuration.  
Docker Container images host the stateless functions responsible for running algorithms
on subpopulations(these can be used directly in a FaaS implementation).
% Need to improve the next paragraph or maybe move it to another place - Mario
Researchers then define the configuration of services used by
a multi-population algorithm using a YAML file. Any researcher can deploy and
start all the services from this configuration file with a single command. Once
the services are up, multiple instances of an experiment can be sent to a queue
for their execution. The platform includes containerized services that are
compatible with other container orchestration technologies like Kubernetes and
Docker Swarm. The services included are a Redis Server used for the Message
Queue implementation and logging and an Experiment Controller implemented in
Python. Developers of sub-population processing functions have the freedom to
implement new population-based algorithms or operators in the language they
choose inside a container, as they only need to subscribe to a channel of the
Message Queue and have the ability to read the population and parameters as JSON  
file.
